---
title: "lab-07-simpsons.Rmd"
author: "Dana Mohammed "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer:
The data is observational as the description states that is based on age, smoking, and mortality, which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)
```

Your answer;
There are 1314 observations. As we know every row is an observation.

3.  

```{r number_variables}
names(Whickham)
```

Your answer:  
There are 3 variables, "outcome", "smoker", and "age"

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)
```

Your answer: 
Using the `unique()` function on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data

one of the best ways to visualise categorical data is through the use of bar charts.
```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +
geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +
geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +geom_boxplot()
```


4. 

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```
The abbreviation of this study is that fewer smokers died than non-smokers
Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
 502+230
 230/732
```
smoker no  (732) :   31,4 (dead) >> (68,6) alive
smoker yes (582) :   23,8 (dead) >> (76,2) alive
The abbreviation of this study is that fewer smokers died than non-smokers
6.

```{r age_cat}
Whickham <- Whickham%>% mutate (age_cat = case_when (age <= 44 ~ "18-44", age > 44. & age <= 64 ~ "45-64", age > 64 ~ "65+"))
```

7. 

```{r plot_smokeVSincome_3}
ggplot(data=Whickham, aes(x=outcome, y=smoker,color=smoker, fill=age_cat)) + geom_bar(stat="identity") +facet_wrap(~age_cat)
```

Knit, commit, and push to github.